TO FIND

### 2013-05-17 ms.tex
Line wrapping

%========================================================%

\section*{Abstract}
Background: \\
Question: \\
Approach: \\
Key results: \\
Conclusion\\

\textbf{Keywords:} metacommunity ecology, neutral theory, species
sorting, patch dynamics, spatial networks, dispersal\\

@@

%========================================================%
\section{Model description}
% DG

General introduction to the section

\subsection{Patch dynamics}
blabla

\subsection{Neutral dynamics}

\subsection{Species sorting and mass effect}
blabla

%For each model:
%	- description of the dynamics
%	- description of the algorithm
%	- main predictions
blabla

%========================================================%
\section{Results}
% TP

\subsection{Alpha diversity}

\subsection{Beta diversity}

@@

Our results point to an important, although often overlooked, consequence of
neutral dynamics. Emerging properties of a community (i.e. the number of
species, but also their spatial distribution and co-occurence at different local
patches), are affected by the dispersal structure.

@@

%========================================================%
\printbibliography
\newpage

>>>>

@@

%========================================================%
\section*{Model description}

In this section we describe three toy models representing different
perspectives of metacommunity ecology: patch dynamics, neutral dynamics and
species sorting. While the neutral model is interesting in itself, it is by
comparison with a model without any interactions (patch dynamics) and with niche
differentiation (species sorting) that we will be able to fully understand the
interaction between these processes and landscape structure. Despite neutral,
competitive interactions in neutral models are very strong because of the
zero-sum assumption (the community is always at carrying capacity). We will
first review the fundamental assumptions of each model with their description
(Table 1 summarizes the parameters and variables that are used), and then
briefly discuss their main predictions. Simulation results are presented in the
next section, with the corresponding R code provided in the Supplementary
Material.

\begin{table*}
	\centering
	\begin{tabular}{llccc}
		\hline
		 & Definition & Patch dynamics & Neutral & Species-sorting \\
		\hline
		\textbf{Variables} &     &    &    &   \\
		$p$ & Occupancy  & X &    &   \\
		$N$ & Local population size &    & X & X \\
		$Z$ & Local rel. abund. &  & X & X \\
		$P$ & Rel. abund. in the neighborhood &  & X & X \\
		$s$ & Local species richness & X & X & X \\
		$d$ & Node degree & X & X & X \\
		$C$ & Prob. of a colonization event & X &  &  \\
		$I$ & Prob. of a colonization event & X &  &  \\
		$Pr$ & Recruitment prob. &  & X & X \\
		$\lambda$ & Survival prob. &  &  & X \\

		\textbf{Indices} & & & &\\
		$x,y$ & Node location & X & X & X \\
		$i,j$ & Species & X & X & X\\
		$n$ & Microsite & & & X \\

		\textbf{Parameters} & & & &\\
		$S$ & Size of regional species pool & X & X & X \\
		$c$ & Colonization prob. & X & & \\
		$e$ & Extinction prob. & X & & \\
		$J$ & Local carrying capacity & & X & X \\
		$m$ & immigration prob. from neigh. & & X & X \\
		$M$ & immigration prob. from metaco. & & X & X \\
		$k$ & Death prob. & & X & X \\
		$u$ & Niche optimum & & & X \\
		$b$ & Niche breadth & & & X \\
		$E$ & Microsite env. conditions & & & X \\
		$\overline{E}$ & Local env. average & & & X \\
		$\sigma$ & Local env. variance & & & X \\
		$\overline{E_R}$ & Regioal env. average & & & X \\
		$\sigma_R$ & Regional env. variance & & & X \\
		\hline
	\end{tabular}
	\caption{List of variables, indices and parameters from the three models}
\end{table*}

%------------------------
\subsection*{Patch dynamics}

The simplest metacommunity model is a $S$ species extension of traditional
metapopulation models \parencite(Hanski1999). The standard Levins metapopulation
model \parencite(Levins1969) describes the stochastic colonizations and
extinctions of population over a homogenous landscape. The basic unit is the
population. The Levins model tracts the dynamics of occupancy (the fraction of
the landscape that is occupied) with an ordinary differential equation and
therefore assumes an infinite landscape. The simulation model we run is more
realistic as it simulates a finite number $N$ of discrete patches. The rules
described in the previous section were used to generate connectivity matrices
along four scenarios (Fig. 1): global dispersal (connected graph), a lattice, a
random geometric graph and a random tree graph. A patch $x$ shares $d_x$ links
with neigbhouring patches (its degree). At each time step (the simulation model
is discrete in time), the probability that a colonist coming from an occupied
patch $y$ arrives at patch $x$ is $cd_y^{-1}$, where $c$ is the probability a
colonization event takes place if all connected patches are occupied. The
expected probability that a colonist arrives to patch $x$ from patch $y$ is then
$C_{ixy}=cp_{iy}d_y^{-1}$, where $p_{i}y$ is the probability that patch $y$ is
occupied by species $i$. The probability that an extinction occurs in a given
patch is $e$. The Levins model is for a single species, but a basic
metacommunity patch dynamics model could be run by aggregating $S$ independent
metapopulation models \parencite(Hanski1997). There are no interactions in this
simple model, which means there is no limit to local species richness and no
carrying capacity. Competitive, mutualistic and predator-prey interactions have
been added to this framework (e.g.
\textcite{Tilman1994,Klausmeier1998,Holt1996}) but we will keep this model
minimal for the sake of comparison with the neutral model.

Predictions of the patch dynamics metacommunity model are quite straightforward.
First, a fundamental result of metapopulation ecology is that persistence will
occur if colonization probability is larger than extinction probability ($c>e$).
Given that all species are the same, then we should expect the regional
diversity to be $S$ if this condition is satisfied and $0$ if not. The situation
is however more complex in spatially explicit landscapes with complex
connectivity matrices \parencite{Hanski1998}. Spatially explicit dispersal
usually reduces the occupancy and thereby the likelihood of persistence. The
second prediction is that, given spatial variation in connectivity, there will
be spatial variation in occurrence probability. Given the above formulation of a
colonization event to occur, the probability that an empty location is colonized
is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation basically tells us that the
colonization probability will increase asymptotically with the degree of a patch
(because of the product). It is easy to show from metapopulation theory that the
occurrence probability in a patch is then $p_ix=I_x(I_ix+e)^{-1}$. The feedback
between local and regional dynamics arises because all $p_ix$ from the landscape
are dependent from each other. Simulation are usually conducted to solve the
model for a large landscape, but numerical solutions are theoretically possible.
The aggregation across the $S$ species of the regional species pool is obtained
by taking the summation of occurrence probabilities over all species, $s_x =
\sum{p_i}$. Because in this model all species are equal, we expect the local
species richness to be a linear function of the patch degree. Multi-species
analysis of metapopulation models also reveals interesting predictions on other
aspects of community organization at various spatial scales such as species-area
relationship \parencite{Hanski1997}, and proved to be useful in conservation
ecology with predictions of extinctions following habitat destruction
\parencite{Nee1994,Rybicki2013}.

%------------------------
\subsection*{Neutral dynamics}

Neutral theory introduces strong competitive interactions by assuming there is a
finite number of individuals that could occupy a  patch. There are different
ways to simulate this \emph{zero-sum rule} \parencite{Bell2000,Hubbell2001}, but
they all result in the same constraint that the increase in abundance of a
species could only occur after an equivalent decrease by another species. One
important change in the formulation of most neutral models relative the patch
dynamics model presented above is therefore that it is individual-based, not
population based. We therefore considered in our toy model of neutral dynamics
that each local patch is made of $J_x$ individuals. The model tracts the local
abundance of all species $N_{ix}$ in each local patch. At each time step an
individidual dies with probability $k$. Recruitment only occurs in vacant sites,
similarly to a tree by tree replacement process in a closed canopy forest.

The formulation of the recruitment probability is the central piece of all
neutral models, making possible the coupling with the metacommunity and
neighbouring patches. We adopt a simple formulation in our model based on
\parencite{Gravel2006}. The approach is conceptually similar to placing a seed
trap in a gap in a forest and picking a seed at random among the one that fall
in to determine the identity of the recruited species. The composition of the
seed pool in that trap will be a mixture of local dispersal and immigrants from
the metacommunity. For simplicity, we consider three spatial scales of dispersal
but it would be easy to generalize the approach to a continuous seed dispersal
kernel \parencite{Gravel2006}. The parameter $m$ is the probability that the
recruit is a migrant from neighbouring patches, $M$ is the probability it comes
from a larger (and fixed) metacommunity, and consequently by substraction
$1-m-M$ is the probability it comes from local dispersal. The fraction
$N_{ix}J_x^{-1}$ is the local relative abundance and $P_ix$ is the relative
abundance of species $i$ in the seed pool coming from neigbouring patches $x$.
The relative abundance in the neighborhood is weighted by the degree of the
connected nodes because some nodes will spread their seeds across a higher
number of nodes and thus contribute less to the seed pool. We consider simply
$P_{ix} =\frac{\sum d_y^{-1}P_{iy}}{\sum dy^{-1}}$.  We assume for simplicity
(and without loss of generality, \parencite{Bell2000}) that the relative
abundance in the metacommunity is uniform, ie equal to $S^{-1}$. This
immigration prevents the collapse of the metacommunity because otherwise all
species except one will face extinction by ecological drift (speciation prevents
this phenomenon to occur in \textcite{Hubbell2001}). The local recruitment
probability is consequently $Pr_{ix} = MS^{-1} + mP_{ix} +
(1-m-M)N_{ix}J_x^{-1}$.

The model is neutral because it assumes that the probabilities of a local
recruitment, an immigration and a mortality event are all equal. Demographic
stochasticity is the source of variations in abundance but larger disturbances
could be simulated as well, as long as they hit all species with the same
probability, independently of their density. The fundamental feature of neutral
dynamics is therefore the ecological drift, defined as population changes
emerging from neutraly stable population dynamics. It can be measured as the
variance between replicated time series of community dynamics
\parencite{Gravel2011}. \textcite{Hubbell2001} provides a very comprehensive
analysis of the model, with specific attention to the effect of the different
parameters on the drift (and consequently variance in abundance) and time to
extinction. Despite its simplicity, the neutral model is surprisingly rich in
the predictions it makes. \textcite{Bell2001} and \textcite{Hubbell2001}
analyzed the ability of neutral models to predict species abundance
distributions, range-abundance relationship, spatial variation in abundance,
species-area relationship, community turnover (beta-diversity) and
co-occurrence. Other than the ecological equivalence assumption, one of the most
criticized aspect of neutral models is the realism of the speciation process and
the required speciation rates to sustain species richness
\parencite{Ricklefs2003,Etienne2007}. Recent neutral models with more credible
speciation models \parencite{Rosindell2009,Desjardins2012a} revealed the
difficulty to maintain diversity in neutral models over macro-evolutionary time
scales. These models however also generated interesting novel predictions on
endemic species richness \parencite{Rosindell2011,Desjardins2012b}.

%------------------------
\subsection*{Species-sorting and mass effect}

The species-sorting and the mass effect perspectives build on the notion of
species-specific responses to a spatially varying environment
\parencite{Leibold2004a}. There are various ways to simulate such dynamics and
we picked the lottery models, in line with tradition \parencite{Mouquet2002} and
for its proximity to the neutral model described above. Competition for space
occurs during recruitment after the death of an adult. The recruitment is a
lottery against potential candidates. The recruitment probability is however
biased by species specific response to local environmental conditions.

The lottery dynamics described above for the neutral model assume there is a
very large number of offsprings that are candidate for recruitment but only one
will survive and develop to the adult stage. The effect of a differentiation to
local environmental conditions could be implemented at this stage.The $J_x$
individuals all experience unique environmental conditions $E_{nx}$ called a
microsite $n$. We considered a patch average $\overline{E_x}$, with a
within-patch variance $\sigma_x$. The regional average is $\overline{E_R}$ and
the regional variance $\sigma_R$ (for simplicity we considered normal
distributions, but different distributions will lead to different regional
similarity constraints \parencite{Mouquet2003,Tilman2004,Gravel2006}). We
consider that a fraction $\lambda_{inx}$ of offsprings reaching the microsite
where recruitment occurs will survive. The recruitment probability is therefore
biased because only some species will be able to cope with the local
environmental conditions. For tractability we define the relative abundance in
the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. The
calculation of the relative abundance in the seed rain is the same as the
neutral model but the recruitment probability differs because only a fraction of
offspring survive. It is formulated as $Pr_{ix} =
\frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing
the relationship between microsite conditions and survival could take various
forms and we used the traditional gaussian curve describing the niche,
$\lambda_{inx} =  \exp{-\frac{(E_{nx}-u_i)^2}{2\Pi b_i^2}}$, where $u_i$ is the
niche optimum and $b_i$ is niche breadth. Note that the model will converge to a
neutral model when the niche breadth tends to infinity (which is in fact how we
simulated neutral dynamics in the Supplementary Material to minimize the
complexity of the code).

Analyses of similar models with a combination of dispersal and species-sorting
shown that predictions are extremely variables and depend on the distribution of
environmental conditions, niche optimums and breadth. For instance, a
well-studied prediction of neutral models is the species abundance distribution.
It was shown that niche models can predict similar distributions
\parencite{Tilman2004,Gravel2006}. The main prediction is nonetheless that
stable and predictable (meaning which species will coexist) if species are
sufficiently dissimilar. Local species richness will first depend on the joint
effects of local heterogeneity and niche breadth because coexistence requires a
limiting similarity \parencite{Schwilk2005}. Local species richness could be
increased by a mass effect when dispersal is consistently supplying individuals
coming from more favorable locations (refuges). The limiting similarity required
to maintain regional coexistence depends on the amount of dispersal because
exchanges among communities homogenizes environmental conditions. This is one of
the main result from the species sorting theory and a clever example of
local-regional feedbacks: increasing dispersal promotes local coexistence, but
on the other hand it diminishes regional coexistence. Only the best average
competitors will remain at very high dispersal.  We therefore expect a
hump-shaped relationship between dispersal and alpha diversity, with a peak at
intermediate dispersal. On the other hand, we expect a monotonic decrease of
beta and gamma diversity with dispersal \parencite{Mouquet2003}. This prediction
has been validated in some experiments \parencite{Venail2008, Logue2011}.

\section{Results}

In fig. 1, we illustrate the outcome of simulations using the neutral model on
the four different network shapes considered in this article. All of these
networks have the same number of nodes (e.g. spatial sampling sites), but both
different number of edges (e.g. dispersal routes between sampling sites) and
patterns of connectivity between nodes. We ask how these differences in
connectivity will shape the emerging properties of the community under the
scenarios represented by each meta-community model.

In fig. 2,

@@

Our results point to an important, although often overlooked, consequence of
neutral dynamics. Emerging properties of a community (i.e. the number of
species, but also their spatial distribution and co-occurence in different local
patches), are affected by the dispersal structure.

@@

%========================================================%
\printbibliography

\newpage
%========================================================%
\section*{Figure legends}

%------------------------
\subsection*{Figure 1}
\textbf{Illustration of the four simulated landscapes}. The color code
represents the local species richness simulated with a neutral model, ranked
from the poorest (red) to the richest (blue). Parameters are: $P = 100$, $N =
25$, $m = 0.2$, $M = 0.01$, $k = 0.1$, $J_x = 100$.

%------------------------
\subsection*{Figure 2}
\textbf{Relationship between local species richness and metrics of node
centrality}. The upper two panels are simulation results conducted with the
random geometric graph illustrated at Fig. 1 and the lower two panels are runs
with the random tree graph.

%------------------------
\subsection*{Figure 3}
\textbf{Bray curtis dissimilarity as a function of geographic distance}.

%------------------------
\subsection*{Figure 4}
\textbf{Bray curtis dissimilarity as a function of topological distance}.

\newpage
%========================================================%

\section*{Figures}

%------------------------
\subsection*{Figure 1}

\begin{figure}[ht!]
	\centering\includegraphics[width=0.3\textwidth]{Networks.pdf}
\end{figure}

%------------------------
\subsection*{Figure 2}

\begin{figure}[ht!]
	\centering\includegraphics[width=0.75\textwidth]{Centrality.pdf}
\end{figure}

%------------------------
\subsection*{Figure 3}

\begin{figure}[ht!]
	\centering\includegraphics[width=0.65\textwidth]{BetaGeoDist.pdf}
\end{figure}

%------------------------
\subsection*{Figure 4}

\begin{figure}[ht!]
	\centering\includegraphics[width=0.65\textwidth]{BetaTopoDist.pdf}
\end{figure}
